### 2015-12-03 extras/paper/Paper_ADC_Sigma-Delta_en_FPGA.tex
Agregado Paper y algunas simulaciones

%% bare_conf.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[a4paper,conference]{IEEEtran}
% Add the compsoc option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%

\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  \graphicspath{{./eps/} }%{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.eps} %.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\usepackage[tight,footnotesize]{subfigure}
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{ADC $\Sigma\Delta$ en FPGA}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{
\IEEEauthorblockN{Nahuel Carducci, Andr\'es Demski, Ariel Kukulanski, e Iv\'an Paunovic}
\IEEEauthorblockA{Estudiantes de la carrera Ingenier\'ia Electr\'onica,\\Facultad Regional Buenos Aires, Universidad Tecnol\'ogica Nacional}
}
%\and
%\IEEEauthorblockN{Homer Simpson}
%\IEEEauthorblockA{Twentieth Century Fox\\
%Springfield, USA\\
%Email: homer@thesimpsons.com}
%\and
%\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
%\IEEEauthorblockA{Starfleet Academy\\
%San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212\\
%Fax: (888) 555--1212}}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

\begin{abstract}
%\boldmath
Se realiza un ADC en una FPGA, usando una topología Sigma-Delta para mantener al mínimo el hardware externo, y aprovechando los recursos de la FPGA. Con esto se logra evitar el uso de un ADC externo, y tener una forma sencilla de implementar uno a bajo costo.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introducci\'on}
% no \IEEEPARstart
Un ADC sigma delta est\'a basado en los conceptos de \textit{Oversampling} y \textit{Noise Shaping} (distribuci\'on del ruido). El \textit{Oversampling} consiste en muestrear a una frecuencia mayor que la de Nyquist, por lo que el ruido de cuantizaci\'on pasa a distribuirse en un ancho de banda m\'as grande, y al pasar luego por un filtro pasa bajos se lo disminuye. El \textit{Noise Shaping} consiste en redistribuir el ruido de cuantizaci\'on de forma tal que quede fuera de la banda de inter\'es, logr\'andose una reducción mayor del ruido de cuantizaci\'on al pasar por un filtro pasabajos (ver Fig.1).

\begin{figure}[!b]
\centering
\includegraphics[width=2in]{Noise_Shaping.eps}
\caption{Arriba Oversampling. Abajo Oversampling+Noise Shapping}
\label{fig_1}
\end{figure}
La topolog\'ia de un ADC sigma delta consiste b\'asicamente en el diagrama en bloques de la Fig.2. Se modela al cuantizador del ADC dentro del lazo de realimentaci\'on como ruido aditivo blanco. Las transferencias son:
\begin{align}
H_x(z)=\frac{H(z)}{1+H(z)} \qquad
H_n(z)=\frac{1}{1+H(z)}
\end{align}
donde $H_x$ y $H_n$ son las transferencias para la señal y para el ruido respectivamente. Siendo H un filtro pasabajos, Hx también será un filtro pasabajos (mientras sea H en la banda de paso mayor a 1) y Hn un filtro pasa altos. De esta forma se obtiene \textit{Noise Shaping}.
\begin{figure}[!t]
\centering
\includegraphics[height=1.8in]{Sigma-Delta_Tiempo_Discreto.eps}
\caption{Diag. bloques modulador $\Sigma\Delta$}
\label{fig_2}
\end{figure}

Para el desarrollo del proyecto se uso el kit de desarrollo Mojo v3 el cual posee una fpga Spartan6.
%\hfill mds
%\hfill January 11, 2007

\subsection{Modulador $\Sigma\Delta$ en tiempo discreto}
Basado en el diagrama en bloques anterior, una modulador primer orden consiste en un acumulador:
\begin{align}
H(z)=\frac{z^{-1}}{1-z^{-1}}
\end{align}
De esto se obtiene que:
\begin{align}
H_x(z)=z^{-1} \qquad
H_n(z)=1-z^{-1}
\end{align}
La transferencia para la señal es simplemente un retardo, y la del ruido es una diferencia.
Para implementar un filtro de \'orden k se suele recurrir al esquema:
\begin{align}
H_n(z)=\left( 1-z^{-1}\right)^{k}
\end{align}

\subsection{Ruido de cuantizaci\'on}
Al suponer el ruido blanco uniformemente distribuido entre -$\Delta$/2 y $\Delta$/2, siendo $\Delta$ el alto del escal\'on al cuantizar se tiene que:

\begin{align}
S_q(f) &= \frac{\Delta^{2}}{12f_s}\\
P_{q}(f) &=\int\limits_{-f_s/2}^{f_s/2}S_q(f) \, |H_n(f) \times H_D(f)|^{2} df
\end{align}

Donde $S_q,P_q y H_D$ son la densidad espectral de potencia de ruido, la potencia de ruido de cuantizaci\'on a la salida, y la transferencia del filtro digital respectivamente.
Teniendo en cuenta un filtro digital ideal con corte en $f_p$, y un modulador de orden k se obtiene que la relaci\'on se\~nal a ruido de cuantizaci\'on en funci\'on del \textit{oversampling ratio}
\footnote[1]{
Se lo obtiene teniendo en cuenta que en el integrando para $OSR\gg 1$ es $f\leq f_p\ll f_s$ y $|H_n(e^{j2 \pi \frac{f}{f_s}})|^{2}= \sin(\pi \frac{f}{f_s})^{2k} \approx (\pi\frac{f}{f_s})^{2k}$ y que la potencia un seno de plena escala es $P_{sin}=\frac{1}{2} [(2^N-1)\frac{\Delta}{2}]^{2}$
}:

\begin{align}
SQNR &\approx 1.76+20\,log_{10}(2N-1)+10\,(2k+1)\,log_{10}OSR\nonumber\\
&\quad
+10\,log_{10}(2k+1)-20k\,log_{10}\pi \\
OSR&=\frac{fp}{fs/2}
\end{align}
En la Fig.3 se ilustra el resultado.
\begin{figure}[!t]
\centering
\includegraphics[height=1.8in]{SQNR_ideal_segun_OSR}
%\def\svgwidth{2.5in}
%\input{Noise_Shaping2.eps}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
\caption{SQNR en funci\'on OSR para diferentes ordenes}
\label{fig_3}
\end{figure}
%\subsubsection{Subsubsection Heading Here}
%Subsubsection text here.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals/conferences use top floats
% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
% command of the stfloats package.

\section{Modulador $\Sigma\Delta$ en tiempo continuo}
En la siguiente imagen se presenta un sigma delta de tiempo continuo, donde se ve reemplazado el cuantizador por su modelo de ruido aditivo:
Las transferencias para la se\~nal y para el ruido quedan como:
\begin{align}
H_x(e^{sT_s}) &= \frac{H(s)}{1+H_{eq}(e^{sT_s})} \\
H_n(z) &= \frac{1}{1+H_{eq}(z)} \\
h_{eq}[n] &= (h\ast h_{DAC})(nT_s)
\end{align}
Donde la respuesta de lazo equivalente en tiempo discreto $H_{eq}(z)$ se la obtiene por la transformaci\'on de invarianza al impulso \cite{DSP:Pro-Man} de $H(z)H_{DAC}(z)$ dada por (10). Para un retenedor de orden cero queda como:
\begin{align}
 H_{DAC}(s)&=\frac{1-e^{-sT_s}}{s}\\
 Z^{-1}(\frac{H_eq(z)}{1-z^{-1}})[n]&=L^{-1}(\frac{H(s)}{s})(nT_s)
\end{align}
Partiendo del modulador de primer orden en tiempo discreto y con la anterior transformaci\'on se obtiene:
\begin{align}
 H(s)&=\frac{1}{sT_s}\\
 H_x(e^{sT_s})&= \frac{1-e^{-sT_s}}{s}
\end{align}
Donde el filtro antes del cuantizador es un integrador, y la trasferencia para la se\~nal es un pasabajos con ceros en multiplos de $f_s$, lo cual reduce notablemente el alias en la banda de paso.
\subsection{Implementaci\'on}

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{Sigma-Delta_Tiempo_Continuo}
\caption{Diag. bloques $\Sigma\Delta$ tiempo continuo}
\label{fig_4}
\end{figure}

\begin{figure}[!b]
\centering
\includegraphics[height=0.9in]{Respuesta_BP_Modulador_Combinada}
%\caption{Respuesta modulador en la banda de paso}
%\label{fig_5}
%\centering
%\includegraphics[width=2.5in]{Respuesta_Ruido_BP_Modulador}
%\caption{Respuesta ruido en la banda de paso}
%\label{fig_6}
\centering
\includegraphics[height=0.9in]{Respuesta_Modulador_Completa}
\caption{Arriba: Respuesta banda de paso de la se\~nal (derecha), y ruido (izquierda). Abajo: Respuesta se\~nal completa}
\label{fig_5}
\end{figure}
La topolog\'ia m\'as sencilla para la implementaci\'on es el modulador RC. La cuantificaci\'on es realizada por una entrada LVDS de la FPGA utilizada com\'unmente para comunicaci\'ones diferenciales y posteriormente la discretizaci\'on temporal es realizada por un flip-flop. La tensi\'on de entrada va en el rango de 0v a la alimentación de la fpga (3.3v), y la referencia está a mitad de escala (1.65v).  Haciendo un breve an\'alisis del circuito, se ve que la tensi\'on $V_{int}$ tomar\'a valores cercanos a la referencia, y como respecto a \'esta la amplitud de la salida ser\'a 1.65v, el comparador presenta una ganancia para la se\~nal. Esto es lo que lo hace posible funcionar, ya que el filtro debe tener ganancia mayor a uno. Generalmente se la estima basado en que cuando no hay entrada, la salida Vq oscilar\'a entre el estado bajo y alto en cada clk, y la ganancia del comparador ser\'ia .
\begin{align}
G=\frac{1}{|H(j\pi fs)|}
\end{align}
Realmente, está correlacionada con Vin y es un efecto que produce alinealidades. La transferencia del filtro RC es:
\begin{align}
H(s)=\frac{1}{2+sCR}
\end{align}
La transferencia a lazo abierto queda como G*H(s). 
Se obtiene la transferencia de lazo equivalente como:
\begin{align}
H_{eq}(z)=\frac{1}{2}\frac{1-e^{-\frac{2}{RC}T_s}}{1-e^{-\frac{2}{RC}T_s}z^{-1}}
\end{align}
Al PLL de la fpga se ajust\'o a 45.3125MHz, y al sistema digital se lo hizo decimar 2048 veces, por lo que a la salida se tendrán 21.13 Ksps. El valor de resistencia usada fue de 150k, y el capacitor de 470p (corte en 4.3kHz).
Con esto se puede graficar las transferencias para la señal y para el ruido con las fórmulas antes indicadas (Fig.5). En las im\'agenes se ve primero la transferencia se\~nal banda de paso, donde el ripple es despreciable. En la segunda se ve la transferencia para el ruido en la banda de paso. Con esto se puede estimar en forma te\'orica el ENOB realizando una integraci\'on num\'erica, dando como resultado 13.6. En la \'ultima imagen se ve el cero transferencia de se\~nal en fs, lo que reduce el alias en la banda de paso.

\subsection{Simulaci\'on}

Se simul\'o un comparador, incluy\'endose efectos de hist\'eresis y ruido sobre la referencia. En la simulaci\'on no se incluyeron las caracter\'isticas el\'ectricas de los buffers de entrada y salida de la fpga. En la Fig.6 se ve una simulaci\'on de la respuesta para una señal de 1kHz.

\begin{figure}[!t]
\centering
\includegraphics[height=2in]{Salida_modulador}
\caption{Arriba: Entrada, Medio: Respuesta a salida modulador, Abajo: Espectro de la salida donde se visualiza el \textit{Noise Shaping}}
\label{fig_6}
\end{figure}

\begin{figure}[!b]
\centering
\includegraphics[width=3.5in]{CIC_Topologia}
\caption{Estructura de un filtro CIC}
\label{fig_7}
\end{figure}

\section{Filtro decimador CIC}
El filtro CIC (Cascaded Integrator-Comb) presentado por Hogenauer en 1981,  es un filtro pasa bajos que tiene como principal caracter\'istica su eficiente implementaci\'on, logrando bajar la frecuencia de trabajo con un filtro sencillo que presenta poco alias.

Se puede observar que cada par acumulador (a la izquierda decimador), y \textit{comb} (a la derecha), forman un filtro de media m\'ovil con ventana de ancho R*M. La transferencia de \'este es:
\begin{align}
H_{CIC}(z)= \left( H_I(z) H_c(z)\right)& ^{N}=\left( \frac{1-z^{-RM}}{1-z^{-1}}\right) ^N\\
\left| H_{CIC}(e^{j2\pi\frac{f}{R}})\right| &=\left| \frac{\sin(\pi M\,f)}{\sin(\pi \frac{f}{R})}\right| ^{N}
\end{align}
Siendo f normalizada a la downsampling rate.
Se observa que los ceros del CIC se producen en los m\'ultiplos de la low sampling rate, es decir, las zonas de alias. El m\'aximo alias se da en f=1-fp, y se puede tener s\'olo en cuenta este valor ya que el aporte de los siguientes l\'obulos es mucho menor.

\subsection{Implementaci\'on y simulaci\'on}
El filtro CIC se determin\'o de 6 etapas, para el cual se utilizaron 12 sumadores de 55 bits. Para reducir el tiempo combinacional y poder tener una frecuencia de trabajo mayor, cada etapa fue registrada. A su vez, se implement\'o, entre las etapas, el bloque de decimaci\'on con un factor de 512. El retraso en los \textit{comb} se determin\'o de M=1.  Las salidas del CIC son almacenadas en una Dual-Port BlockRam la cual tiene la caracter\'istica de poder realizar la lectura de dos datos de manera simultanea. En la Fig.8 se ve la respuesta en la banda de paso, y la primer zona de alias.

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{Respuesta_CIC}
\caption{Respuesta CIC. Arriba: Banda de paso, Abajo: Primer zona de alias}
\label{fig_8}
\end{figure}

\section{Filtro FIR}

\begin{figure}[!b]
\centering
\includegraphics[width=3.5in]{FIR_Topologia}
\caption{Topolog\'ia usada en implementaci\'on FIR}
\label{fig_9}
\end{figure}

El filtro FIR busca ecualizar la transferencia en la banda de paso del CIC (la cual no es plana), y eliminar el ruido de cuantizaci\'on fuera de esta antes de la decimaci\'on final. Se lo dise\~n\'o por el m\'etodo de sampleo y ventaneo con una ventana Kaiser, donde en la banda de paso se samplea la inversa del filtro CIC usado y fuera de esta ceros. 
Para la implementación del FIR, se utiliz\'o la estructura de PreAdder-Multiplicador-Acumulador que posee el bloque DSP de la FPGA. Se utiliz\'o un filtro FIR sim\'etrico de 256 coeficientes  con frecuencia de corte normalizada de 0.22. Al ser sim\'etrico, s\'olo es necesario almacenar la mitad de los coeficientes y tambi\'en se reduce a la mitad el c\'alculo al poder utilizar el preadder.  Posteriormente, se realiza una \'ultima decimaci\'on por un factor de 4, dando como resultante una decimaci\'on total de 2048.
En la Fig.9 se ve que el filtro FIR tienen una atenuaci\'on de

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{Respuesta_FIR}
\caption{Respuesta FIR}
\label{fig_10}
\end{figure}

\begin{table}[!b]
%% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{Resultado Simulaci\'on ENOB}
\label{Tabla_1}
\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
\begin{tabular}{|c|c|c|}
\hline
Frec. entrada & Frec. Fiteo & ENOB\\
\hline
1k & 999.99 & 10.66\\
\hline
2k & 2000.00 & 10.84\\
\hline
4k & 4000.00 & 10.74\\
\hline
8k & 8000.00 & 10.24\\
\hline
9k & 8999.99 & 62.90\\
\hline
10k & 9999.98 & 8.6\\
\hline
\end{tabular}
\end{table}

\section{Simulaci\'on}
Se simul\'o el modulador descrito previamente, as\'i como el filtro CIC y el FIR en Matlab.
Durante las simulaciones, se realizaron pruebas de linealidad barriendo la entrada con una rampa de muy baja frecuencia y se pudo ver que el sistema posee un comportamiento totalmente lineal.
Tambi\'en se realizaron simulaciones barriendo en frecuencia para obtener la cantidad efectiva de bits.
A los datos obtenidos se les aplic\'o un \textit{fitting} con una senoidal de tres grados de libertad (Amplitud, frecuencia y fase). Con esto obtuvimos la se\~nal que mejor aproxima a los puntos obtenidos. Luego se calcula la potencia de la se\~al de error. El SINAD (Relaci\'on Se\~nal a ruido + distorsi\'on) se lo estima como raz\'on entre la potencia de la se\~nal obtenida en la regresi\'on y de la de error. El ENOB (cantidad de bits efectivos) se lo define a partir de la relaci\'on se\~nal a ruido de cuantizaci\'on de un ADC de N bits.
\begin{align}
SINAD=6.02 \times ENOB + 1.73
\end{align}
Los resultados obtenidos en la simulaci\'on se los resume en la Tabla1.

\begin{figure}[!t]
\centering
\includegraphics[width=2.5in]{Simulacion_fiteo}
\caption{\textit{Fitting} de la senoidal}
\label{fig_11}
\end{figure}

\section{Mediciones}

\begin{figure}[!t]
\centering
\includegraphics[width=2.5in]{Medicion_Linealidad}
\caption{A la izquierda con la correcci\'on linealidad, a la derecha sin. En ambos casos de arriba a abajo: respuesta entrada a salida y \textit{fitting}, desvio estandar, error en el \textit{fitting}}
\label{fig_12}
\end{figure}

Para obtener los resultados de las mediciones se realiz\'o una comunicaci\'on UART entre la FPGA y la PC. El set de medici\'on const\'o de un generador de funciones Agilent 3352A, controlado v\'ia ETHERNET desde la computadora. Luego se analizaron los datos medidos mediante scripts realizados en Matlab$\textregistered$.
En la medici\'on de linealidad se obtuvo una distorsi\'on en valores de entrada cercanos a la referencia de tensi\'on. La linealidad fue corregida digitalmente, aunque de momento la corrección fue s\'olo implementada en Matlab$\textregistered$ y no en la FPGA.

Luego se realiz\'o el c\'alculo de la cantidad efectiva de bits mediante el mismo m\'etodo que el usado en la simulaci\'on. En la imagen se puede comparar el resultado para el ADC linealizado digitalmente y sin linealizar.
La presencia de arm\'onicos disminuye el ENOB cuando no se lo linealiza (curva roja Fig.13 Fig.14). Al subir en frecuencia, los arm\'onicos quedan fuera de la banda de paso y son filtrados, por lo que el ENOB sube. Al linealizar (curva azul), mejora el ENOB a bajas frecuencias, ya que en ellas la correcci\'on agrega distorisi\'on que no hab\'ia.

El THD se calcul\'o sumando la potencia de las arm\'onicas. En el gr\'afico se ve c\'omo la THD disminuye al linealizar. A bajas frecuencias y sin linealizar, result\'o de 0.003$\%$ , equivalente a -45dB. Teniendo en cuenta que la SINAD era de 45dB se ve que \'esta estaba dominada por el contenido arm\'onico.

\begin{figure}[!t]
\centering
\includegraphics[width=2.5in]{Medicion_ENOB}
\caption{ENOB seg\'un la frecuencia}
\label{fig_13}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=2.5in]{Medicion_THD}
\caption{THD seg\'un la frecuencia}
\label{fig_14}
\end{figure}

\section{Conclusi\'ones}
La realización obtenida del proyecto logr\'o cumplir parcialmente los objetivos. Se logr\'o una implementaci\'on sencilla y de bajo costo, ocupando bajos recursos en la fpga (ver tabla2). De las \textit{block RAM} de 16kB no se uso toda la memoria, pero solo pueden ser divididas en bloques de 8kB, la que se uso como ROM le quedo un bloque disponible. El mayor problema en la implementaci\'on fue la distorsi\'on en el modulador. Los algoritmos de simulaci\'on y bancos de prueba usados, facilitaran la implementaci\'on de posibles mejoras.
\begin{table}
\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{Recursos FPGA usado}
\label{Tabla_2}
\centering
\begin{tabular}{|c|c|c|c|}
\hline
Bloque & Usados & Disponibles & $\%$ \\
\hline
Slices FF & 1078 & 11440 & 9\\
\hline
Slices Luts & 1009 & 5720 & 17\\
\hline
Block RAM(16kB) & 2 & 32 & 6\\
\hline
DSP48A1 & 1 & 16 & 6\\
\hline
\end{tabular}
\end{table}

% conference papers do not normally have an appendix

% use section* for acknowledgement
\section*{Acknowledgment}

The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, %3rd~ed.\hskip 1em plus
%	0.5em minus 0.4em\relax Harlow, England: %Addison-Wesley, 1999.
\bibitem{DSP:Pro-Man}
J.G.~Proakis and D.G.~Manolakis \emph{Digital Signal Processing Principles, Algorithms, and Aplicattions}, 3rd~ed.\hskip 1em plus
	0.5em minus 0.4em\relax Patrice-Hall, 1996.

\end{thebibliography}

% that's all folks
\end{document}