### 2016-04-10 extras/paper_en_ingles/Paper_ADC_Sigma-Delta_en_FPGA.tex
Comenzando traducción paper

%% bare_conf.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[a4paper,conference]{IEEEtran}
% Add the compsoc option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

%% TILDES
%\usepackage[utf8]{inputenc}
%\usepackage[spanish]{babel}

\renewcommand{\tablename}{Tabla}

% *** GRAPHICS RELATED PACKAGES ***
%

\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  \graphicspath{{../paper/eps/} }%{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.eps} %.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  \graphicspath{{./eps/}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\usepackage[tight,footnotesize]{subfigure}
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Design, Implementation and Characterization of a Sigma-Delta Analog-to-Digital Converter in FPGA}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{
\IEEEauthorblockN{Carducci, Nahuel \\ Demski, Andr\'es\\  Kukulanski, Ariel\\  Paunovic, Iv\'an }
\IEEEauthorblockA{Departamento de Ingenier\'ia Electr\'onica,\\Universidad Tecnol\'ogica Nacional, Facultad Regional Buenos Aires.}
}
%\and
%\IEEEauthorblockN{Homer Simpson}
%\IEEEauthorblockA{Twentieth Century Fox\\
%Springfield, USA\\
%Email: homer@thesimpsons.com}
%\and
%\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
%\IEEEauthorblockA{Starfleet Academy\\
%San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212\\
%Fax: (888) 555--1212}}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

\begin{abstract}
%\boldmath
While analog to digital converters are often part of microcontrollers, this doesn't happen with FPGA, forcing a developer to obtain an external converter for his projects. As a FPGA is a widely used tool for DSP, we developed a Sigma-Delta ADC in the FPGA, using very low external hardware. In this publication we describe the design, implementation and measure of this converter.

\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% no \IEEEPARstart

Sigma-delta ($\Sigma\Delta$) analog to digital converter is based on two main conceps: \textit{Oversampling} and \textit{Noise Shaping}.
\textit{Oversampling} consists on sampling a signal with a frequency higher than Nyquist's. This distributes the quantization noise in a bigger bandwidth.
\textit{Noise shaping} consists on redistributing the quantization noise in such a way that most of it will be out of the band of interest.
Both concepts are shown in Fig.~\ref{fig:NS}.
The modulator is followed by a digital low pass decimation filter, eliminating most of the in-band quantization noise. This allows converters with high ENOB.

\begin{figure}[!b]
\centering
\includegraphics[width=1.7in]{Noise_Shaping.eps}
\caption{Up: Oversampling. Low: Oversampling+Noise Shapping}
\label{fig:NS}
\end{figure}

The $\Sigma\Delta$ ADC topology is in the block diagram of Fig. ~\ref{fig:SDTD}. The quantization noise is modeled as additive white noise ($N(z)$). The transfers functions for the signal (STF) and the noise (NTF) are:

\begin{align}
STF(z)=\frac{H(z)}{1+H(z)} \qquad
NTF(z)=\frac{1}{1+H(z)}
\end{align}

Being H a low pass filter, STF will be a low pass filter, and NTF will be a high pass filter, obtaining the previously mentioned Noise Shaping.

\begin{figure}[!t]
\centering
\includegraphics[height=1in]{Sigma-Delta_Tiempo_Discreto.eps}
\caption{Block diagram of the $\Sigma\Delta$}
\label{fig:SDTD}
\end{figure}

% % Poner en materiales y metodos
For the develop of the project, we used the developing kit \textit{Mojo v3}, that incorpores a \textit{Xilinx Spartan 6 FPGA}.
% %

In ~\ref{fig:DBADC} we show the basic blocks of the converter, where we can distinguish the modulator (analog stage), and two digital filters (CIC and FIR). The CIC (Cascaded integrator–comb) is a decimation filter that can be implemented with few%low
FPGA resources. The decimation factor we used is 512. The FIR is the one that will eliminate the noise accumulated in the high frequencies. After the FIR, we decimate again with a factor of 4, obtaining a total decimation factor of 2048.

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{Diagrama_Sistema_ADC.eps}
\caption{Block diagram of the ADC}
\label{fig:DBADC}
\end{figure}

\subsection{Discrete-Time $\Sigma\Delta$ Modulator}

If H in Fig. 2 as an accumulator:

\begin{align}
H(z)=\frac{z^{-1}}{1-z^{-1}}
\end{align}

Using Eq. 1, we obtain the STF and NTF as:

\begin{align}
STF(z)=z^{-1} \qquad
NTF(z)=1-z^{-1}
\end{align}

Then the STF is just a delay, while the NTF is a high pass filter as is wanted.
To implement a k-order filter, the following NTF is used:

\begin{align}
NTF(z)=\left(1-z^{-1}\right)^{k}
\end{align}

\subsection{Quantization noise}

Assuming the quantization noise (QN) is white and uniformly distributed between -$\Delta/2$ and $\Delta/2$, being $\Delta$ the quantization step, it can be shown that:

% % poner una referencia

\begin{align}
S_q(f) &= \frac{\Delta^{2}}{12f_s}\\
P_{q}(f) &=\int\limits_{-f_s/2}^{f_s/2}S_q(f) \, |NTF(f) \times H_D(f)|^{2} df
\end{align}

where $S_q$ is the QN power spectral density and $P_q$ is the QN power after Noise Shapping and Digital Filtering, being $H_D$ the transfer function of the digital filter.
For an ideal digital filter with cutoff frequency FP, and a modulator of order k, the signal to quantization noise relationship is:

\begin{align}
SQNR &\approx 1.76+20\,log_{10}(2N-1)+10\,(2k+1)\,log_{10}OSR\nonumber\\
&\quad
+10\,log_{10}(2k+1)-20k\,log_{10}\pi \\
OSR&=\frac{fp}{fs/2}
\end{align}

where OSR is the oversampling ratio.

% %CAMBIAR FORMULA DE ARRIBA POR LA FORMULA APROXIMADA QUE APARECE SIEMPRE CON UNA REFERENCIA Y SIN ESTA EXPLICACION
%\footnote[1]
%{
%Se lo obtiene teniendo en cuenta que en el integrando para $OSR\gg 1$ es $f\leq f_p\ll f_s$ y $|H_n(e^{j2 \pi \frac{f}{f_s}})|^{2}= \sin(\pi \frac{f}{f_s})^{2k} \approx (\pi\frac{f}{f_s})^{2k}$ y que la potencia un seno de plena escala es $P_{sin}=\frac{1}{2} [(2^N-1)\frac{\Delta}{2}]^{2}$.
%Usualmente aparece la formula con $2^{N}-1\approx2^N$.
%} es:

The resulting SQNR is plotted in Fig.~\ref{fig:SQNR} as a function of the OSR for different filters orders.

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{SQNR_ideal_segun_OSR}
%\def\svgwidth{2.5in}
%\input{Noise_Shaping2.eps}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
\caption{SQNR as a function of de OSR for different filters orders}
\label{fig:SQNR}
\end{figure}

\section{Continuous-Time $\Sigma\Delta$ Modulator}

Unlike the discrete-time $\Sigma\Delta$ modulator, here the sampling is made inside the modulator, right after quantization (Fig.~\ref{fig:SDTC}). The STF and NTF are obtained as:

\begin{align}
STF(e^{sT_s}) &= \frac{H(s)}{1+H_{eq}(e^{sT_s})} \\
NTF(z) &= \frac{1}{1+H_{eq}(z)} \\
h_{eq}[n] &= (h\ast h_{DAC})(nT_s)
\end{align}

where the equivalent discrete-time $H_eq$ response is obtained with the impulse-invariance transform\cite{DSP:Pro-Man} of $H(s)H_{DAC}(s)$ defined in (11).
Using the DAC as a Zero-order hold is:

\begin{align}
 H_{DAC}(s)&=\frac{1-e^{-sT_s}}{s}\\
 Z^{-1}(\frac{H_{eq}(z)}{1-z^{-1}})[n]&=L^{-1}(\frac{H(s)}{s})(nT_s)
\end{align}

With the first order discrete time modulator and the previous transformation, we obtain:

\begin{align}
 H(s)&=\frac{1}{sT_s}\\
 STF(e^{sT_s})&= \frac{1-e^{-sT_s}}{s}
\end{align}

Being H an integrator, and the STF a low pass filter with zeros in multiples of $f_s$, which reduces the alias in the passband.

\begin{figure}[!t]
\centering
\includegraphics[height=1in]{Sigma-Delta_Tiempo_Continuo}
\caption{Block diagram of a continuous-time $\Sigma\Delta$ modulator}
\label{fig:SDTC}
\end{figure}

% %Seguir desde aca
%A. Implementation
% %The simplest implementation for the modulator is an RC filter (fig. 6). The quantization is done by a LVDS input of the FPGA [6], commonly used for differnetial communications, and the time discretization is done by a flip flop.

\subsection{Implementaci\'on}

\begin{figure}[!t]
\centering
\includegraphics[width=3in]{Sigma-Delta_RC}
\caption{Circuito utilizado para el modulador $\Sigma\Delta$}
\label{fig_SDRC}
\end{figure}

La topolog\'ia m\'as sencilla para la implementaci\'on es el modulador RC (Fig.\ref{fig_SDRC}). La cuantificaci\'on es realizada por una entrada LVDS de la FPGA\cite{Sp6-IO} utilizada com\'unmente para comunicaciones diferenciales y posteriormente la discretizaci\'on temporal es realizada por un flip-flop. La tensi\'on de entrada posee un rango de 0v a la alimentación de la fpga (3.3v), y la referencia está a mitad de escala (1.65v).  Haciendo un breve an\'alisis del circuito, se ve que la tensi\'on en la entrada del comparador tomar\'a valores cercanos a la referencia, y como respecto a \'esta la amplitud de la salida ser\'a 1.65v, el comparador presenta una ganancia para la se\~nal. Esto es lo que hace posible funcionar al modulador pasivo, ya que el filtro debe tener ganancia mayor a uno. Generalmente a esta ganancia se la estima pasivando la entrada de señal, lo cual provoca una oscilación de la mitad de la frecuencia de \textit{clk}. La ganancia del comparador ser\'ia :
\begin{align}
G(s)=\frac{1}{|H(j\pi f_s)|}
\end{align}
Realmente, esta ganancia está correlacionada con Vin provocando alinealidades en el sistema.\\
La transferencia del filtro RC es:
\begin{align}
H(s)=\frac{1}{2+sCR}
\end{align}
La transferencia a lazo abierto por lo tanto es $G*H(s)$. 
Se obtiene la transferencia de lazo equivalente como:
\begin{align}
H_{eq}(z)=\frac{z^{-1}}{2}\frac{1-e^{-\frac{2}{RC}T_s}}{1-e^{-\frac{2}{RC}T_s}z^{-1}}
\end{align}
El PLL de la fpga fue ajustado a 45.3125MHz, y al sistema digital se lo hizo decimar 2048 veces, por lo que a la salida se tendrán 22.13 Ksps. El valor de resistencia usada fue de 150k, y el capacitor de 470pF (corte en 4.3kHz).
Con esto se puede graficar las transferencias para la señal y para el ruido con las fórmulas antes indicadas (Fig.~\ref{fig:Resp_mod}). Observando la transferencia para la señal, se ve que el ripple es despreciable. Luego con la transferencia para el ruido en la banda de paso, se estimó el ENOB realizando una integraci\'on num\'erica que dio como resultado 13,6. Por último, se observa que la primer zona de alias tiene un m\'aximo de -72dB.

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{Respuesta_Modulador_Combinada}
\caption{Arriba: Transferencia para la se\~nal en la banda de paso. Abajo: Transferencia para el ruido (izquierda) y primer zona de alias para la se\~nal}
\label{fig:Resp_mod}
\end{figure}

\subsection{Simulaci\'on}

Para realizar la simulación del modulador, se modelizó el comparador incluyendo efectos de histéresis y adición de ruido sobre la tensión de referencia. En la simulaci\'on no se incluyeron las caracter\'isticas el\'ectricas de los buffers de entrada y salida de la \textit{FPGA}. En la Fig.~\ref{fig:Sim_mod} se ve una simulaci\'on de la respuesta para una señal de 1kHz. En esta se ve que la salida del integrador RC tiene una oscilación pico a pico 2mV, lo que significa una ganancia de señal G=1650. Al resimular el ENOB del modulador RC con esta ganancia se obtiene un resultado de 10.93 (el alias se mantiene en -72dB).

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{Simulacion_Modulador}
\caption{De arriba hacia abajo: Señal de entrada, señal en la entrada del comparador, señal de salida del modulador,  espectro de la salida donde se visualiza el \textit{Noise Shaping} (izquierda), zoom en la banda de paso (derecha).}
\label{fig:Sim_mod}
\end{figure}

\section{Filtro decimador CIC}

\begin{figure}[!b]
	\centering
	\includegraphics[width=3.5in]{CIC_Topologia}
	\caption{Estructura de un filtro CIC}
	\label{fig:CIC_top}
\end{figure}

El filtro CIC (Cascaded Integrator-Comb) presentado por Hogenauer en 1981\cite{CIC:Hog},  es un filtro pasa bajos que tiene como principal caracter\'istica su eficiente implementaci\'on (ver Fig.~\ref{fig:CIC_top}), logrando bajar la frecuencia de trabajo con un filtro sencillo que presenta poco alias en el ancho de banda de interés.

Se puede observar que cada par de acumulador (a la izquierda del decimador), y \textit{comb} (a la derecha), forman un filtro de media m\'ovil con ventana de ancho R*M. La transferencia de \'este es:

\begin{align}
H_{CIC}(z)= \left( H_I(z) H_c(z)\right)& ^{N}=\left( \frac{1-z^{-RM}}{1-z^{-1}}\right) ^N\\
\left| H_{CIC}(e^{j2\pi\frac{f}{R}})\right| &=\left| \frac{\sin(\pi M\,f)}{\sin(\pi \frac{f}{R})}\right| ^{N}
\end{align}

siendo f la frecuencia normalizada al downsampling rate.\\

Se observa que los ceros del CIC se producen en los m\'ultiplos de la low sampling rate, es decir, las zonas de alias. El m\'aximo alias es en f=1-fp (normalizado), y se puede tener s\'olo en cuenta este valor ya que el aporte de los siguientes l\'obulos es mucho menor, considerando fp como el ancho de banda del filtro ideal que va a continuación del filtro CIC.

\subsection{Implementaci\'on y simulaci\'on}
El filtro CIC se determin\'o de 6 etapas, para el cual se utilizaron 12 sumadores de 55 bits. Para reducir el tiempo combinacional y poder tener una frecuencia de trabajo mayor, cada etapa fue registrada. A su vez, se implement\'o entre las etapas el bloque de decimaci\'on con un factor de 512. El retraso en los \textit{comb} se determin\'o M=1. Las salidas del CIC son almacenadas en una Dual-Port BlockRam\cite{Sp6-BR} la cual tiene la caracter\'istica de poder realizar la lectura de dos datos de manera simultánea. Los datos almacenados en la misma fueron truncados a 16 bits para optimizar la síntesis. En la Fig.~\ref{fig:CIC_resp} se muestra la respuesta en la banda de paso, y la primer zona de alias. Se puede observar que hay una atenuación de 1.5dB en la banda de paso, y el alias que se producirá es de -100dB.

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{Respuesta_CIC}
\caption{Respuesta del CIC: Banda de paso (Arriba) y primer zona de alias (Abajo)}
\label{fig:CIC_resp}
\end{figure}

\section{Filtro FIR}

El filtro FIR busca ecualizar la transferencia en la banda de paso del CIC (la cual no es plana), y eliminar el ruido de cuantizaci\'on fuera de la banda de interés antes de la 
decimaci\'on final. Se lo dise\~n\'o por el m\'etodo de sampleo y ventaneo con una ventana Kaiser, donde en la banda de paso se muestrea la inversa de la respuesta en frecuencia del filtro CIC usado y fuera de esta se determina con ceros\cite{FIR_Comp}. En la Fig.~\ref{fig:FIR_res} se muestra la respuesta del filtro FIR diseñado, que posee una atenuaci\'on de 80dB.
Para la implementación del FIR, se utiliz\'o la estructura de PreAdder-Multiplicador-Acumulador (Fig.~\ref{fig:FIR_top}) que posee los bloques \textit{DSP} de la \textit{FPGA}\cite{Sp6-DSP}. Se utiliz\'o un filtro FIR sim\'etrico de 256 coeficientes  con frecuencia de corte normalizada de 0.22. Al ser sim\'etrico, s\'olo es necesario almacenar la mitad de los coeficientes y tambi\'en se reduce a la mitad el c\'alculo al poder utilizar el preadder.  Posteriormente, se realiza una \'ultima decimaci\'on por un factor de 4, dando como resultante una decimaci\'on total de 2048. El resultado final fue truncado a 16 bits.

\newpage

\begin{figure}[!t]
	\centering
	\includegraphics[width=3.5in]{Respuesta_FIR}
	\caption{Respuesta del FIR}
	\label{fig:FIR_res}
\centering
\includegraphics[width=3.5in]{FIR_Topologia}
\caption{Topolog\'ia usada en la implementaci\'on del FIR}
\label{fig:FIR_top}
\end{figure}

\begin{table}[!b]
	%% increase table row spacing, adjust to taste
	\renewcommand{\arraystretch}{1.3}
	% if using array.sty, it might be a good idea to tweak the value of
	% \extrarowheight as needed to properly center the text within the cells
	\caption{Resultado Simulaci\'on ENOB}
	\label{Tabla:sim}
	\centering
	%% Some packages, such as MDW tools, offer better commands for making tables
	%% than the plain LaTeX2e tabular which is used here.
	\begin{tabular}{|c|c|c|}
		\hline
		Frec. entrada & Frec. Fiteo & ENOB\\
		\hline
		1k & 999.99 & 11.17\\
		\hline
		2k & 2000.00 & 11.01\\
		\hline
		4k & 4000.00 & 10.68\\
		\hline
		8k & 8000.00 & 9.84\\
		\hline
		9k & 8999.99 & 9.56\\
		\hline
		10k & 9999.98 & 7.85\\
		\hline
	\end{tabular}
\end{table}

\section{Simulaci\'on}
Se simul\'o el modulador descripto previamente, as\'i como el filtro CIC y el FIR en Matlab\textregistered.
Durante las simulaciones, se realizaron pruebas de linealidad barriendo la entrada con una rampa de muy baja frecuencia y se pudo ver que el sistema posee un comportamiento totalmente lineal.
Tambi\'en se realizaron simulaciones barriendo en frecuencia para obtener la cantidad efectiva de bits. Sabiendo que los mismos estan definidos por: 
\begin{align}
ENOB = \dfrac{SINAD-1.73}{6.02}
\end{align}
se procedió a buscar un estimador para el SINAD (Razón entre la potencia de la Se\~nal y el ruido en conjunto con la distorsi\'on). Para ello, se excitó el sistema con una señal senoidal y se almacenaron los datos de la salida de la simulación. Posteriormente, se les realizó un \textit{fitting} con una función senoidal con tres grados de libertad (Amplitud, frecuencia y fase). Con esto, se obtuvo la función que mejor la aproxima la señal obtneida y se realizó una resta punto a punto para así, poder determinar la potencia del ruido y con ello, el SINAD. Con esto, se obtuvieron los bits efectivos resumidos en la Tabla~\ref{Tabla:sim}.

\section{Mediciones}

\begin{figure}[!b]
\centering
\includegraphics[width=3.5in]{Linealidad}
\caption{Arriba: Curva de linealidad (azul) y regresión lineal (roja). Medio: Desvío Estándar de las mediciones. Abajo: Error entre la curva de linealidad y regresión lineal}

\label{fig:lin}
	\centering
	\includegraphics[width=3.5in]{Linealidad_codo}
	\caption{Ampliación del codo en la curva de linealidad. Curva de linealidad (azul), regresión lineal (rojo)}

	\label{fig:lincodo}
\end{figure}

Para realizar las mediciones se implementó una comunicaci\'on UART entre la \textit{FPGA} y la PC a 921600 baudios. El set de medici\'on const\'o de un generador de funciones Agilent 3352A\cite{AgilentWG}, controlado v\'ia TCP desde la computadora. La adquisición de datos, su almacenamiento y posterior análisis se realizó mediante la herramienta de cálculo Matlab\textregistered. \\Se llevaron a cabo tres mediciones: Linealidad, THD y ENOB.\\
Para realizar la medición de linealidad, se procedió a barrer todo el rango de tensión (0 a 3.3v) con un paso de 3mV obteniendo 1000 muestras por cada uno. Posteriormente, se realizó una regresión lineal de las muestras obteniendo las curvas representadas en la figura \ref{fig:lin}. Como se aprecia en la figura \ref{fig:lincodo}, la curva presenta una alinealidad significativa. Para análisis posteriores, se procedió a generar una curva de corrección para contrarrestar este efecto. Con tal fin, se generó una interpolación de la señal de error para poder tener $2^{16}$ valores para corregir.\\
Para el cálculo de THD y ENOB, se procedió a excitar el sistema con señales senoidales \textit{full-range} y se hizo un barrido en frecuencia entre 0 y 11KHz con pasos logarítmicos para los cuales se adquirieron un mínimo de 1000 muestras y con la restricción de tener un número entero de períodos de la señal senoidal, así evitar futuros problemas.
Una vez adquiridas las muestras, mediante el uso del periodograma con ventana de Kaiser, se realizaron las mediciones del THD. Para ello, conociendo la frecuencia de la fundamental (con la cual se excitó el sistema), se identificaron las armónicas y se determinó su potencia. Las cuales se reemplazaron en la siguiente expresión:
\begin{align}
THD = \dfrac{\sum_{i=1}^{M}P_i }{P_0}\cdot100
\end{align}
Siendo $P_0$ la potencia de la fundamental y $P_i$ la potencia de los respectivos armónicos.\\
Para la medición del ENOB, se realizó el mismo procedimiento llevado acabo en la simulación pero en este caso con datos adquiridos.\\
Estas mismas mediciones (ENOB y THD) se realizaron una segunda vez pero para los datos aplicando previamente la corrección obtenida durante el \textit{test} de linealidad. Como resultado, se obtuvieron las curvas trazadas en las figuras \ref{fig:ENOB} y \ref{fig:THD}, en azul representando las curvas con corrección y en rojo sin ella.
\begin{figure}[!t]
\centering
\includegraphics[height=2in]{Medicion_ENOB}
\caption{ENOB en funcion de la frecuencia. Azul: Corregida. Rojo: Sin Corregir}
\label{fig:ENOB}
\centering
\includegraphics[height=2in]{Medicion_THD}
\caption{THD en función de la frecuencia. Azul: Corregida. Rojo: Sin Corregir}
\label{fig:THD}
\end{figure}

\section{Discusión y Conclusi\'ones}
Analizando las curvas obtenidas, podemos notar que el limitante de nuestro proyecto es la distorsión armónica generada en el modulador. Esta distorsión no fue percibida durante las simulaciones debido a que en la misma no se tuvieron en cuenta las características eléctricas de los componentes y de la \textit{FPGA}. Si se aplica la corrección digital al sistema, los resultados son prometedores en las bajas frecuencias alcanzando los 9 bits efectivos. Para frecuencias mayores a 100 Hz, esta corrección empieza a perder efectividad debido a que se compensan armónicos que ya han sido eliminados por los filtros digitales.\\
La implementación del proyecto logr\'o cumplir los objetivos propuestos. Se logr\'o una implementaci\'on de un conversor analógico digital de 9 bits y 22KHz de frecuencia de muestreo que se puede considerar sencilla y de bajo costo y consumiendo pocos recursos en la \textit{FPGA} (ver Tabla~\ref{Tabla:rescursos}). \\
\begin{table}[!h]
	\renewcommand{\arraystretch}{1.3}
	% if using array.sty, it might be a good idea to tweak the value of
	% \extrarowheight as needed to properly center the text within the cells
	\caption{Recursos FPGA usados}
	\label{Tabla:rescursos}
	\centering
	\begin{tabular}{|c|c|c|c|}
		\hline
		Bloque & Usados & Disponibles & $\%$ \\
		\hline
		Slices FF & 1113 & 11440 & 9\\
		\hline
		Slices Luts & 932 & 5720 & 16\\
		\hline
		Block RAM(16kB) & 2 & 32 & 6\\
		\hline
		DSP48A1 & 1 & 16 & 6\\
		\hline
	\end{tabular}
\end{table}

Los algoritmos de simulaci\'on y bancos de prueba utilizados, facilitarán la implementaci\'on de posibles mejoras en el sistema y la comprobación de su funcionamiento. Además, la implementación en la \textit{FPGA} puede ser reutilizada ya que todos los parámetros del diseño son totalmente configurables, pudiéndose hacer fácilmente pruebas con distintas frecuencias de clock, OSR, parámetros del filtro CIC y coeficientes del FIR. Por lo que a partir de esta primer implementaci\'on se podrá mejorar fácilmente el sistema.
 En próximos desarrollos, se llevará  la corrección digital dentro de la \textit{FPGA}, específicamente luego del CIC. Con esto, se logrará corregir la señal previamente al filtrado con lo cual se espera obtener una mayor planicitud en la curva de ENOB. También, se pretende tener mejores modelos de simulación para poder así identificar el causante de la alinealidad y poder atenuar su influencia. Se simularán e implementarán moduladores de mayor orden para que, con este mismo diseño digital, poder obtener mayor ENOB y, en lo posible, mayor frecuencia de muestreo.

% conference papers do not normally have an appendix

% use section* for acknowledgement
%\section*{Acknowledgment}

%The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, %3rd~ed.\hskip 1em plus
%	0.5em minus 0.4em\relax Harlow, England: %Addison-Wesley, 1999.
\bibitem{DSP:Pro-Man}
J.G. Proakis and D.G. Manolakis \emph{Digital Signal Processing Principles, Algorithms, and Aplicattions}. 3rd ed. Patrice-Hall, 1996, pp 671-676.
\bibitem{Tesis:Hellman}
Hellman J. \emph{Implementation of a Low-Cost Analog-to-Digital Converter for Audio Applications Using an FPGA}, MS Tesis, Dep. Electric. Eng., Linköpings Universitet, Linköping, Suecia, 2013.
\bibitem{S-D_Antialias}
Peev, P.; De Vuyst, B.; Rombouts, P.; Hamoui, A.A. \emph{An anti-aliasing filter inspired by continuous-time $\Sigma\Delta$ modulation.} ICECS, 2008.
\bibitem{CIC:Hog}
	Hogenauer, Eugene. \emph{An economical class of digital filters for decimation and interpolation.} Acoustics, Speech and Signal Processing, IEEE Transactions on 29.2 (1981): 155-162.
\bibitem{FIR_Comp}
	\emph{Understanding CIC Compensation Filters}, App. Note 455 (v1.0). Altera, 2007.
\bibitem{Sp6-IO}
	\emph{Spartan 6 FPGA SelectIO Resources}, UG381 (v1.7). Xilinx, 2015.
\bibitem{Sp6-BR}
	\emph{Spartan 6 FPGA Block RAM Resources}, UG383 (v1.5). Xilinx, 2011.
\bibitem{Sp6-DSP}
	\emph{Spartan 6 FPGA DSP48A1 Slice}, UG389 (v1.2). Xilinx, 2014.
\bibitem{AgilentWG}
	\emph{33521A and 33522A Function / Arbitrary Waveform Generator Data Sheet}. Agilent. 
\end{thebibliography}

% that's all folks
\end{document}